---
title: "Predictive Analysis Project_Group 2"
author: "Phùng Quang Anh, Lê Bảo Trung, Phạm Hoàng Long, Ngô Gia Tráng, Nguyễn Anh Dũng"
date: "`r Sys.Date()`"
output:
  html_document: 
    code_download: true
    # code_folding: hide
    highlight: pygments
    # number_sections: yes
    theme: "flatly"
    toc: TRUE
    toc_float: TRUE
---

# Import dataset
```{r}
#Install and download the necessary packages
library(dplyr)
library(stringr)
library(rio)
library(broom)
library(tidyr)
library(caret)
library(stats)
library(MASS)
library(readxl)  
library(mlbench)
library(ggplot2)
library(psych)
library(magrittr)
library(tree)
```

```{r}
#Import data
training <- read.csv("C:/Users/84974/QuangAnhPhung/Documents/ACTUARY NĂM 4/PA (Predictive Analytics)/Final exam/training_set.csv")
test <- read.csv("C:/Users/84974/QuangAnhPhung/Documents/ACTUARY NĂM 4/PA (Predictive Analytics)/Final exam/test_set.csv")
attach(training)
attach(test)
```


# Question 1: Data preprocessing
## a. name: extract the car brand and make it a new feature in replacement of the original feature
```{r}
training = training %>%
  mutate(name = word(name, 1, sep = " "))
test = test %>%
  mutate(name = word(name, 1, sep = " "))
```

## b.mileage & engine & max_power: keep only the numeric value and get rid of characters
```{r}
training = training %>%
  mutate(mileage = as.numeric(gsub("[^0-9.]", "", mileage)),
         engine = as.numeric(gsub("[^0-9]", "", engine)),
         max_power = as.numeric(gsub("[^0-9.]", "", max_power)))

test = test %>%
  mutate(mileage = as.numeric(gsub("[^0-9.]", "", mileage)),
         engine = as.numeric(gsub("[^0-9]", "", engine)),
         max_power = as.numeric(gsub("[^0-9.]", "", max_power)))

```
## c.torque: create 2 features, one "nm" with Nm values and one "rpm" with rpm values. In case, they are 2 rpm values, calculate the average of these values.
```{r}
#Remove dirty observations which doesn't include Nm or rpm

training = training %>%
  mutate(torque = ifelse(grepl("Nm", torque, ignore.case = TRUE), torque, ""),
         torque = ifelse(grepl("rpm", torque, ignore.case = TRUE), torque, ""))

test = test %>%
  mutate(torque = ifelse(grepl("Nm", torque, ignore.case = TRUE), torque, ""),
         torque = ifelse(grepl("rpm", torque, ignore.case = TRUE), torque, ""))

#Split into 2 columns named "Nm" and "rpm"
training= separate(training, torque, into = c("Nm", "rpm"), sep ="@|/|at", convert =TRUE, extra = "merge" )

test= separate(test, torque, into = c("Nm", "rpm"), sep ="@|/|at", convert =TRUE, extra = "merge" )

#Function for Nm calculation
Nm_convert = function(text) {
  if (is.na(text)){
    result= NA
  } else {
    result = as.numeric(substring(text,1,regexpr("Nm", text,ignore.case = TRUE)-1))
  }
  return(result)
}

# Function for rpm calculation
rpm_convert = function(text) {
  if (is.na(text)){
    result= NA
  } else if (gsub("[0-9]", "", text)=="+/-"){
    result = as.numeric(substring(text,1,regexpr("+/-", text,ignore.case = TRUE)-2))
  } else {
    numbers = as.numeric(unlist(str_extract_all(text, "\\d+")))
    if (length(numbers) == 2) {
      result = mean(numbers)
    } else if (length(numbers) == 1) {
      result = numbers[1]
    } else {
      result = NA
    }
  }
  return(result)
}
#Apply Nm function for the dataset
training = training %>%
  mutate(Nm=sapply(Nm, Nm_convert))

test = test %>%
  mutate(Nm=sapply(Nm, Nm_convert))

#Apply the rpm function to the dataset
training = training %>%
  mutate(rpm=sapply(rpm, rpm_convert))

test = test %>%
  mutate(rpm=sapply(rpm, rpm_convert))
```

## d.Removing rows for which data is missing or dirty.
```{r}
#Remove NA and dirty values
training<-na.omit(training)
test<-na.omit(test)
```

## e. Removing duplicates
```{r}
#Remove duplicate
training<- distinct(training)
test<- distinct(test)

```

# Convert character columns to factor
```{r warning=FALSE}
#Final dataset

factor_type = c("name","fuel","seller_type","transmission","owner","year","seats") 
training[factor_type] =lapply(training[factor_type],as.factor)
test[factor_type] = lapply(test[factor_type],as.factor)

num_type = c("km_driven","selling_price","mileage","engine","max_power","Nm","rpm") 
training[num_type] =lapply(training[num_type],as.numeric)
test[num_type] = lapply(test[num_type],as.numeric)


# Remove observation of seats in test set which are not included in training set
train_seats <- unique(training$seats)
filtered_test_seats <- test %>% anti_join(training, by = "seats")
test <- test %>% filter(!seats %in% filtered_test_seats$seats)
# Final dataset
cleaned_training <- training
cleaned_test <- test
```



# Question 2: Linear regression  
## Build linear regression on the training set 
```{r}

#training a linear regression on the training set
lm <- lm(selling_price ~ ., data = cleaned_training)
```

## a.	Using MSE, compare the training metric with the test metric and conclude about overfitting
```{r}
#Calculate MSE on training set
pred_lm_training = predict(lm, cleaned_training)
mse_lm_training = mean((cleaned_training$selling_price - pred_lm_training)^2)
print(mse_lm_training)

#Calculate MSE on test dataset
pred_lm_test = predict(lm, cleaned_test)
mse_lm_test = mean((cleaned_test$selling_price - pred_lm_test)^2)
print(mse_lm_test)


# Creating dataframe including actual and predicted price
results_test_lm <- data.frame(Actual = cleaned_test$selling_price, Predicted = pred_lm_test)

#Scatter Plot of actual and predicted price
library(ggplot2)

ggplot(results_test_lm, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Actual", y = "Predicted", title = "Predicted vs Actual on Test Set") +
  theme_minimal()
```

## b.Analyze the validity of the model, using a F-test at a p-value’s threshold of 5%
```{r}
#Display model summary and conclude about the P-value of F-test
summary(lm)
```


## c. Analyze the coefficient’s significance, using a t-test at a p-value’s threshold of 5%. Exclude the features that should be excluded  so to build your final model
```{r}
# Set significance threshold (0.05)
#Backward stepwise 
backward_model <- stepAIC(lm, scope = list(upper = lm, lower = ~1), direction = "backward", criterion = "p-value", threshold = 0.05)
# Identify significant variables
selected_variables <- names(coef(backward_model))
print(selected_variables)

#New model
final_lm <- lm(selling_price~.- rpm, data = cleaned_training)
summary(final_lm)
```

## d.	Repeat steps a-b-c for your final model
```{r}
#Calculate MSE on training dataset using final model
pred_lm_training_final = predict(final_lm, cleaned_training)
mean((cleaned_training$selling_price - pred_lm_training_final)^2)

#Calculate MSE on test dataset using final model
pred_lm_test_final = predict(final_lm, cleaned_test)
mean((cleaned_test$selling_price - pred_lm_test_final)^2)

# Creating dataframe including actual and predicted price
results_test_lm_final <- data.frame(Actual = cleaned_test$selling_price, Predicted = pred_lm_test_final)

#Scatter Plot of actual and predicted price
library(ggplot2)

ggplot(results_test_lm_final, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Actual", y = "Predicted", title = "Predicted vs Actual on Test Set") +
  theme_minimal()


```


# Question 3: Elastic net
```{r}
# Extract numeric and non-numeric columns
numeric_columns <- c("selling_price", "km_driven", "mileage", "engine", "max_power", "Nm", "rpm")
non_numeric_columns <- c("name", "fuel", "seller_type", "transmission", "owner")
factor_columns <- c("year", "seats")

# Scale numeric columns
scaled_train <- scale(cleaned_training[, numeric_columns])
scaled_test <- scale(cleaned_test[, numeric_columns])


# Combine scaled numeric columns with non-numeric columns and factor columns
training_scaled <- data.frame(
  cleaned_training[, non_numeric_columns],
  scaled_train,
  cleaned_training[, factor_columns, drop = FALSE]  # Use 'drop = FALSE' to keep factors as a data frame
)

test_scaled <- data.frame(
  cleaned_test[, non_numeric_columns],
  scaled_test,
  cleaned_test[, factor_columns, drop = FALSE]  # Use 'drop = FALSE' to keep factors as a data frame
)
```
## a.	Based on cross-validation (5 folds), check for the best value of the l1-l2 allocation (𝜃) and quantity of regularization (𝜆) and create a model accordingly based on the error 
```{r}
training.control=trainControl(method="repeatedcv",
                           number=5 ,# cross-validation (5 folds)
                           repeats = 5,
                           verboseIter = T)

# training ELastic Net Regression model
elastic_net_model <- train(selling_price ~ ., 
                           training_scaled, 
                           method = "glmnet", 
                           trControl = training.control) 
#Fitting alpha = 0.1, lambda = 0.0139 on full training set

# The best value
print(elastic_net_model$bestTune)
```

## b.	Based on your final model, using MSE, compare the training metric with the test metric and conclude about overfitting 

```{r warning=FALSE}

# Calculate the MSE, performance for training set
pred_elastic_training_scale = predict(elastic_net_model, training_scaled)
pred_elastic_training= pred_elastic_training_scale*sd(cleaned_training$selling_price) + mean(cleaned_training$selling_price)
mean((pred_elastic_training - cleaned_training$selling_price)^2)

# Calculate MSE, performance for test set
pred_elastic_test_scale = predict(elastic_net_model, test_scaled)
pred_elastic_test= pred_elastic_test_scale*sd(cleaned_test$selling_price) + mean(cleaned_test$selling_price)
mean((pred_elastic_test - cleaned_test$selling_price)^2)



# Creating dataframe including actual and predicted price
results_test_elastic <- data.frame(Actual = cleaned_test$selling_price, Predicted = pred_elastic_test)

#Scatter Plot of actual and predicted price
library(ggplot2)

ggplot(results_test_elastic, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Actual", y = "Predicted", title = "Predicted vs Actual on Test Set") +
  theme_minimal()
```

## c. List the features kept in your model as well as their coefficients

```{r}

# Extract variable importance
importance <- varImp(elastic_net_model)
print(importance)
# Extract coefficients
coef <- coef(elastic_net_model$finalModel,s=elastic_net_model$finalModel$lambdaOpt)
print(coef)
```





# Question 4: Decision tree
## Build regression tree model on the training set
```{r}
# Fit a regression tree with target variable "selling_price", all predictors is included
#Tree model
tree_model <- tree(selling_price ~ ., data = cleaned_training)
summary(tree_model)

## Plot the tree
plot(tree_model)
text(tree_model, pretty = 0.5)
```

## a,b.	Using MSE, compare the training metric with the test metric and conclude about overfitting
```{r}
#Calculate the MSE on the training set
pred_tree_training = predict(tree_model, cleaned_training)
mean((pred_tree_training - cleaned_training %>% pull(selling_price))^2)
# Calculate MSE, performance of the test set
pred_tree_test = predict(tree_model, cleaned_test)
mean((pred_tree_test - cleaned_test %>% pull(selling_price))^2)

results_test_tree <- data.frame(Actual = cleaned_test$selling_price, Predicted = pred_tree_test)

#Scatter Plot of actual and predicted price
library(ggplot2)

ggplot(results_test_tree, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Actual", y = "Predicted", title = "Predicted vs Actual on Test Set") +
  theme_minimal()
```

## c.	Based on cross-validation (5 folds), check for the best size of the tree and prune it accordingly
```{r}
# Perform cross-validation to select the optimal tree size
cv_tree <- cv.tree(tree_model)
plot(cv_tree$size, cv_tree$dev, type = "b")
#Identify the tree size with the lowest cross-validated error
#best_tree_size <- which.min(cv_tree$dev)
```

In the cv plot, it can be seen that the pruning tree is the best at 5
```{r}
pruned_tree <- prune.tree(tree_model, best =5 )
#Summary the prunned tree
summary(pruned_tree)
# Plot the prunned tree
plot(pruned_tree) 
text(pruned_tree, pretty = 0.5)
```
#d. Repeat steps a-b on your pruned tree. 
```{r}
## Calculate the MSE on the training dataset of prunned tree
pred_pruned_tree_training = predict(pruned_tree, cleaned_training)
mean((pred_pruned_tree_training - cleaned_training %>% pull(selling_price))^2)
## Calculate the MSE on the test dataset of prunned tree
pred_pruned_tree_test = predict(pruned_tree, newdata = cleaned_test)
mean((pred_pruned_tree_test - cleaned_test$selling_price)^2)

results_test_pruned_tree <- data.frame(Actual = cleaned_test$selling_price, Predicted = pred_pruned_tree_test)

#Scatter Plot of actual and predicted price
library(ggplot2)

ggplot(results_test_pruned_tree, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Actual", y = "Predicted", title = "Predicted vs Actual on Test Set") +
  theme_minimal()
```



# Question 5: Random Forest
## Build the random forest on the traininging set
```{r}
factor_type = c("name","year","fuel","seller_type","transmission","owner","seats") 
cleaned_training[factor_type] =lapply(cleaned_training[factor_type],as.character)
cleaned_test[factor_type] = lapply(cleaned_test[factor_type],as.character)
library(randomForest)
rf_model <- randomForest(selling_price ~ ., data = cleaned_training, ntree = 10, mtry = sqrt(13))
rf_model
```


# a. Using MSE, compare the traininging metric with the test metric and conclude about overfitting
```{r}
pred_rf_training <- predict(rf_model, cleaned_training)
pred_rf_test <- predict(rf_model, cleaned_test)
mse_rf_training <- mean((cleaned_training$selling_price - pred_rf_training)^2)
mse_rf_test <- mean((cleaned_test$selling_price - pred_rf_test)^2)
cat("Training MSE:", mse_rf_training, "\n")
cat("Test MSE:", mse_rf_test, "\n")

# Creating dataframe including actual and predicted price
results_test <- data.frame(Actual = cleaned_test$selling_price, Predicted = pred_rf_test)

# Plot of actual and predicted price
library(ggplot2)

ggplot(results_test, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Actual", y = "Predicted", title = "Predicted vs Actual on Test Set") +
  theme_minimal()
```

# b.	Obtain the feature importance of your model and comment it
```{r}
importance <- importance(rf_model)
print(importance)
```

# c.	Based on cross-validation (5 folds), check for the best number of trees  and model it accordingly
```{r}
library(MASS)
library(caret)
control = trainControl(method = "cv", number = 5, search = "grid")
tuneGrid1 = expand.grid(.mtry = (1:13))
set.seed(1234)
rf_best_mtry <- train (selling_price ~ .,
data = cleaned_training,
method = 'rf',
metric = 'rmse',
tuneGrid = tuneGrid1)
rf_best_mtry
```
RMSE was used to select the optimal model using the smallest value.
The final value used for the model was mtry = 13.

```{r}
best_ntrees = list()
tuneGrid2 = expand.grid(.mtry = 13)
for (ntree in seq(50, 500, 50)) {
set.seed(1234)
rf_best_ntrees = train(selling_price~.,
data = cleaned_training,
method = "rf",
metric = "mse",
tuneGrid = tuneGrid2,
trControl = control,
importance = TRUE,
ntree = ntree)
key = toString(ntree)
best_ntrees[[key]] = rf_best_ntrees
}
rf_tree_results = resamples(best_ntrees)
summary(rf_tree_results)
```
 As we can see, the best number of tree is 200, with the mean value of RMSE is 145430.9 Now I will try
to build a random forest model with mtry = 10 and ntree = 200 to see if there are any improvement.

```{r}
#Build the random forest with new ntree and new mtry
rf_model_new <- randomForest(selling_price ~ ., data = cleaned_training, ntree = 200, mtry = 10)
rf_model_new
```

```{r}
#Calculating MSE
pred_rf_training_new <- predict(rf_model_new, cleaned_training)
pred_rf_test_new <- predict(rf_model_new, cleaned_test)

mse_rf_training_new <- mean((cleaned_training$selling_price - pred_rf_training_new)^2)
mse_rf_test_new <- mean((cleaned_test$selling_price - pred_rf_test_new)^2)
cat("Training MSE:", mse_rf_training_new, "\n")
cat("Test MSE:", mse_rf_test_new, "\n")

# Creating dataframe including actual and predicted price
results_test_new <- data.frame(Actual = cleaned_test$selling_price, Predicted = pred_rf_test_new)

# Plot of actual and predicted price
library(ggplot2)

ggplot(results_test_new, aes(x = Actual, y = Predicted)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(x = "Actual", y = "Predicted", title = "Predicted vs Actual on Test Set (New Model)") +
  theme_minimal()


```

```{r}
importance_new <- importance(rf_model_new)
print(importance_new)
```

